---
title: "MPM02: Group Work"
author: "Authors: Tu Tran, Vanya Todorova, Luca Hüsler"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

## Install and load packages
```{r, message=FALSE}
usePackage <- function(p) 
{
  if (!is.element(p, installed.packages()[,1]))
    install.packages(p, dep = TRUE)
  require(p, character.only = TRUE)
}

packages <- c('plyr', 'tidyverse', 'ggplot2')

for (i in packages){
  usePackage(i)
}
```

## Read files to raw data frame
```{r message=FALSE}

# custom function to add additional column year to the data frame
readcsv <- function(f) {
  flen <- nchar(f)
  fyear <- substring(f, flen-7, flen-4)
  fdata <- read.csv(f, header=TRUE, sep=",")
  fdata <- cbind(year=strtoi(fyear), fdata)
  return (fdata)
}

# define pathes to data
path_electricity <- 'data/dutch_energy/Electricity/'
path_gas <- 'data/dutch_energy/Gas/'

# create list of files to import
files_electricity <- list.files(path = path_electricity, pattern = '*.csv', full.names = T)
files_gas <- list.files(path = path_gas, pattern = '*.csv', full.names = T)

# read and combine all csv-files into a single dataframe with ldply (plyr)
electricity_raw <- ldply(files_electricity, readcsv)
gas_raw <- ldply(files_gas, readcsv)

# check if new column of year added
print(dim(electricity_raw))
head(electricity_raw)

print(dim(gas_raw))
head(gas_raw)

```


## Clean data

Since the data for the provider "enexis" is missing in 2009, we will exclude this year for the analysis:
```{r}
electricity_raw <- electricity_raw[electricity_raw$year != 2009,]
gas_raw <- gas_raw[gas_raw$year != 2009,]

# The year indicated in the CSV-files represents the year of data collection, the actual data is from the year before.
# To be clear about this, we substract 1 from every year
electricity_raw$year <- as.integer(electricity_raw$year -1)
gas_raw$year <- as.integer(gas_raw$year -1)

# Check if column 'year' is correct now in both raw files
table(electricity_raw$year)
table(gas_raw$year)
```

## Part 1
### Data cleaning
```{r}
# check missing values of all column
colMeans(is.na(electricity_raw))
# ok, these 3 columns have more missing values and they are not needed for the analysis -> drop these columns
drops <- c("purchase_area", "street", "city",  "type_conn_perc", "type_of_connection")
electricity_clean <- electricity_raw[, !(names(electricity_raw) %in% drops)]
# check the data frame after dropping
dim(electricity_raw)
dim(electricity_clean)
# check missing values again after dropping
colMeans(is.na(electricity_clean))
# very small amount of delivery_perc has missing values and can be dropped (very limitted data)
electricity_clean <- electricity_clean[!is.na(electricity_clean$delivery_perc),]
# check for distinction of each row (unique observation unit)
distinctdataraw <- distinct(electricity_clean)
nrow(distinctdataraw)
nrow(electricity_clean)

```

### enhance the data with additional aggregated columns
```{r}
electricity_clean <- cbind(electricity_clean, 
                 annual_consume_lowtarif = electricity_clean$annual_consume_lowtarif_perc / 100 * electricity_clean$annual_consume,
                 delivery_give_back = (100 - electricity_clean$delivery_perc) / 100 * electricity_clean$annual_consume,
                 consume_per_connection = (electricity_clean$annual_consume / electricity_clean$num_connections) * (electricity_clean$perc_of_active_connections / 100))

names(electricity_clean)
head(electricity_clean, 3)

```


### Exploring the clean data
```{r}
# check summary (min, max, 1,3 quantiles, median, mean), check data consistency
summary(electricity_clean)
str(electricity_clean)
```

#### Visual overview
Before entering the analysis, it is important to get a feeling for the data and variables of interest. This can be achieved by creating some basic plots.

##### Evolution of total electricity consumption per year
```{r}
electricity_clean %>%
  select(year, annual_consume) %>%
  group_by(year) %>%
  summarise(sum_consumption = sum(annual_consume)) %>%
  ggplot() + 
  geom_col(aes(year, sum_consumption)) +
  labs(title = "Total electricity consumption 2009 - 2018",
           y = "Electricity consumption (kwh)",
           x = "Year")

```
We see that the total electricity consumption is rather stable over the years, even a little bit descending.


##### Evolution of smartmeter percentage per year
```{r}
electricity_clean %>%
  select(year, smartmeter_perc) %>%
  group_by(year) %>%
  summarise(mean_smartmeter = mean(smartmeter_perc)) %>%
  ggplot() + geom_col(aes(year, mean_smartmeter), fill="#880011") +
  labs(title = "Evolution of mean smartmeter percentage 2009 - 2018",
           y = "Mean smartmeter percentage",
           x = "Year")

```
We can see that the mean smartmeter percentage develops in an exponential way. This gives us a first hint: There is a strong correlation between smartmeter percentage and year. This should be considered in the model.

##### Number of active connections
```{r}
electricity_clean %>%
  select(year, num_connections, perc_of_active_connections) %>%
  mutate(num_active_connections = num_connections*perc_of_active_connections/100) %>%
  group_by(year) %>%
  summarise(sum_active_connections = sum(num_active_connections)) %>%
  ggplot() + geom_col(aes(year, sum_active_connections), fill="#880011") +
  labs(title = "Evolution of electricity consumption 2009 - 2018",
           y = "Mean smartmeter percentage",
           x = "Year")


```

##### Evolution of mean consume per connection
```{r}
electricity_clean %>%
  select(year, consume_per_connection) %>%
  group_by(year) %>%
  summarise(mean_consume_per_connection = mean(consume_per_connection)) %>%
  ggplot() + geom_col(aes(year, mean_consume_per_connection), fill="#880011") +
  labs(title = "Evolution of consume per connection 2009 - 2018",
           y = "Mean consume per connection",
           x = "Year")


```

### Prepare train (30%) and test (70%) dataset
```{r}

set.seed(8)
trainindex<- sample(nrow(electricity_clean), 0.3*nrow(electricity_clean))
electricity_clean.train <- electricity_clean[trainindex, ]
electricity_clean.test <- electricity_clean[-trainindex, ]

```

### Part 1.1: Relation between smart meter percentage and total consumption of electricity
#### A: Linear Model
```{r}
# ??? not sure if model based on percentage ok or not?
fit <- lm(delivery_perc ~ smartmeter_perc, data=electricity_clean.train)
trainpred <- predict(fit, electricity_clean.train)
testpred <- predict(fit, electricity_clean.test)
# test error
testerror <- mean((testpred - electricity_clean.test$delivery_perc)^2)
testerror
# train error
trainerror <- mean((trainpred - electricity_clean.train$delivery_perc)^2)
trainerror
# test/train error ratio
ratio_test_train <- testerror/trainerror
ratio_test_train
# R squared
summary <- summary(fit)
summary$r.squared

```

First, we create a linear model $smartmeter\_perc = \beta_0 + \beta_1*annual\_consume$ by using the train dataset that follows.

```{r}

lm.smartmeter <- lm(smartmeter_perc ~ annual_consume + num_connections, data=electricity_clean.train[electricity_clean.train$year == 2018,])
summary(lm.smartmeter)
```

```{r}
plot(electricity_clean.train$smartmeter_perc[electricity_clean.train$year == 2018], electricity_clean.train$annual_consume[electricity_clean.train$year == 2018])
```


#### B: SVM Model

```{r}


```

### Part 1.2:  Relationship between the ratio of low_tarif electricity consumption (low_tarif / total_consumption) and the same predictor

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

## Part 2: Energy efficency
```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```

